PyTracer: Automatically profiling numerical instabilities in Python

### 2021-11-11 paper.tex
Update comments

\documentclass[11pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{todonotes}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage{minted} % Provides listing environment
\usepackage[htt]{hyphenat}
\usepackage{subcaption}
% \usepackage{subfigure}

@@

\newcommand{\Yohan}[1]{\color{green!75!black}\textbf{Yohan:} #1\color{black}\xspace}
\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\pytracer}[0]{PyTracer\xspace}

@@

\begin{abstract}
Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. To address this issue, we developed \pytracer, a profiler to quantify numerical instability in Python applications. \pytracer automatically instruments Python code to produce numerical traces and visualize them interactively in a Plotly dashboard.
\pytracer is designed to work with any type of numerical noise model, including Monte-Carlo Arithmetic, random rounding or random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. \tristan{Add one sentence to summarize the findings} Finally, \pytracer slows down executions by only $3.5 \times$ on average.
\end{abstract}

@@

Dynamic Binary Instrumentation (DBI) operates directly on executables, without the need for recompilation or manual changes. Therefore, it is applicable to any programming language.
CraftHPC~\cite{lam2013dynamic} uses DBI to detect catastrophic cancellations at runtime.
Verrou~\cite{fevotte2016verrou} is a Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts. FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by using shadow computations with higher precision.
Herbgring~\cite{sanchez2017finding} is a Valgrind-based tool to detect
numerical instabilities. It uses a shadow memory to detect precision losses by comparison with results obtained at higher precision. It is combined with symbolic computation to backtrack the root of the error.
We can note that all methods based on a high-precision oracle suppose that computing with a large number of bits is sufficient to obtain an accurate result, which is not always true (see for example, the famous Muller's sequence~\cite{chatelain2018veritracer} \tristan{it looks like your veritracer paper is the first time where the ``famous Muller's sequence" was described, is that so?}).

@@

Three main approaches exist to detect numerical instabilities: stochastic arithmetic, uncertainty or sensitivity analysis, and random seed analysis. \tristan{add key references for uq/sa if any} \pytracer focuses on stochastic arithmetic \tristan{due to...}
Stochastic arithmetic leverages randomness to estimate numerical instabilities coming from the use of floating-point representations. The main idea is to treat round-off errors 
% \tristan{only round-off errors or all numerical errors?} 
as random variables and characterize them statistically.

@@

\pytracer's \texttt{loader} class first loads the original module, then copies it as a new instance of the \texttt{ModuleType} class. It then calls the appropriate instrumentation function for each attribute depending on its type (function, class, or instances --- see below \tristan{iterators?}) and replaces the original module in the \texttt{sys.modules} map with the instrumented version. Finally, it updates all existing references to the original module, such as those in \texttt{\_\_globals\_\_} function attributes containing references to internal function objects \tristan{this sentence doesn't make sense, please update}. As a result, the application will transparently call the instrumented module.

@@

    Returns:
        outputs: Outputs of the wrapped function
    """
    fid, fmodule, fname = info
    function = original_function_cache.id_dict[fid]
    bind = Binding(function, *args, **kwargs)
    stack = self.backtrace()
    time = elements()
    self.inputs(time=time,
                module_name=fmodule,
                function_name=fname,
                function=function,
                args=bind.arguments,
                backtrace=stack)
    outputs = function(*bind.args, **bind.kwargs)
    self.outputs(time=time,
                 module_name=fmodule,
                 function_name=fname,
                 function=function,
                 args=outputs,
                 backtrace=stack)
    return outputs
\end{lstlisting}
    \caption{\pytracer's generic wrapper function that dumps
    the inputs and outputs of the wrapped function.
    \pytracer binds arguments to the function signature to avoid
    mispositioning \tristan{what does mispositioning means here?} and to show readable names in the visualization.
    }
    \label{fig:generic_wrapper}
\end{listing}

@@

We evaluated \pytracer on two widely-used Python libraries: SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference machine-learning framework that
 implements a wide range of state-of-the-art algorithms. For each library, we executed the software tests provided in their respective GitHub repositories. We traced each test five times, having activated MCA in the Python interpreter, BLAS, LAPACK, NumPy, SciPy, and scikit-learn. We repeated the experiment with two MCA modes, Random Rounding (RR) and Full MCA, resulting in 2 experiments for each test. We used the default virtual precision in Verificarlo, namely 24 bits for single-precision floating-point values and 53 bits for double-precision values, which simulates machine error. \tristan{Report scipy and sklearn versions used.}

@@

Tables~\ref{tab:pytracer_scipy_results_summary} and~\ref{tab:pytracer_sklearn_results_summary} summarize 
\pytracer's outcome on SciPy and scikit-learn tests.
Each table reports on the tracing and trace aggregation steps with three possible outcomes: green for success, red for numerical error in the application, and orange for errors in the fuzzy noise injection \tristan{Explain what this is, in a setence here or with a subsubsection if it requires more}. Overall, the tracing and parsing of SciPy and scikit-learn tests in Random Rounding mode was entirely successful for 32/40 tests. The remaining tests failed due to numerical errors in the libraries, which we discuss hereafter. The instrumentation with Full MCA is more invasive and reduced the number of successful executions to 12/40.

\begin{table}[]
    \centering
    \begin{tabular}{|ll|c|c|c|c|}
    \hline
    \multicolumn{2}{|c}{ \multirow{2}{*}{Application} } & \multicolumn{2}{|c|}{RR} & \multicolumn{2}{c|}{Full MCA} \\
    \cline{3-6}
    & & trace & parse & trace & parse  \\
    \hline
    %        &               RR                 &                 MCA               \\
    %        &  numpy         &   no numpy      &      numpy      &     no numpy    \\
    %        & run   & parse  &  run   &  parse &  run   & parse  & run    & parse  \\

    \multicolumn{1}{|c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} } }
    & discrete\_cosine\_transform &  \valid & \valid & \warn & \cross \\ 
    \multicolumn{1}{|c|}{} & fft\_1D & \valid & \valid& \warn & \cross \\
    \multicolumn{1}{|c|}{} & fft\_2D & \valid & \valid & \valid & \valid \\
    \hline 
    \multicolumn{1}{|c|}{  \multirow{5}{2em}{ \rotatebox{90}{Interpolation} } }
    & interpolation\_1D & \valid & \valid & \warn & \cross  \\
    \multicolumn{1}{|c|}{} & multivariate\_data & \valid & \valid & \warn & \cross \\
     \multicolumn{1}{|c|}{} & bspline  & \valid & \valid & \valid & \valid \\
     \multicolumn{1}{|c|}{} & spline\_1D & \valid & \cross & \warn & \cross \\
     \multicolumn{1}{|c|}{} & spline\_2D & \valid & \valid & \warn & \cross \\
    \hline
     \multicolumn{1}{|c|}{ \multirow{12}{2em}{ \rotatebox{90}{Optimization} } }
    & broyden\_fletcher\_goldfarb\_shanno & \valid & \valid & \valid & \valid \\
     \multicolumn{1}{|c|}{} & global\_optimization & \valid & \cross  & \warn & \cross \\
    \multicolumn{1}{|c|}{}  & slsqp & \valid & \valid  & \valid & \valid \\
     \multicolumn{1}{|c|}{} & least\_square\_minimization & \valid & \valid  & \warn & \cross \\
     \multicolumn{1}{|c|}{} & nelder\_mead\_simplex  & \valid & \valid  & \valid & \valid \\
     \multicolumn{1}{|c|}{} & newton\_conjugate\_gradient  & \valid & \valid  & \valid & \cross \\
     \multicolumn{1}{|c|}{} & root\_findings  & \valid & \cross & \warn & \cross \\
    \multicolumn{1}{|c|}{}  & root\_finding\_large & \valid & \cross  & \valid & \cross \\
     \multicolumn{1}{|c|}{} & root\_finding\_large\_preconditionned & \valid & \cross & \valid & \cross \\
    \multicolumn{1}{|c|}{}  & trust\_region\_exact& \valid & \valid  & \valid & \valid \\
     \multicolumn{1}{|c|}{} & trust\_region\_truncated\_generalized\_lanczos & \valid & \valid & \valid & \valid \\
    \multicolumn{1}{|c|}{}  & trust\_region\_newton\_conjugate\_gradient & \valid & \valid & \valid & \valid \\
    \hline
    \end{tabular}
    \caption{\pytracer execution summary on SciPy tests.\\RR: Random Rounding. trace: outcome of \pytracer tracing, parse: outcome of \pytracer trace parsing and aggregation.  \valid: successful run, \warn: error in fuzzy noise injection, \cross: numerical error or divergent traces.}
    \label{tab:pytracer_scipy_results_summary}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \multirow{2}{5em}{Application} & \multicolumn{2}{c|}{RR} & \multicolumn{2}{c|}{MCA} \\\cline{2-5}
                                & trace & parse &  trace & parse \\
    \hline
    %        &               RR                 &                 MCA               \\
    %        &  numpy         &   no numpy      &      numpy      &     no numpy    \\
    %        & run   & parse  &  run   &  parse &  run   & parse  & run    & parse  \\
    Adaboost & \valid & \valid  & \warn & \cross \\
    Bayesian Ridge Regression  & \valid & \valid  & \cross & \cross \\
    Online classifier comparison & \valid & \valid  & \warn & \cross  \\
    K-Means Clustering  & \valid & \valid  & \valid & \cross \\
    Covariance estimation  & \valid & \cross  & \warn & \cross \\
    Decision Tree Regression & \valid & \valid  & \valid & \valid \\
    Digits Classification  & \valid & \valid  & \warn & \cross \\
    Face Recognition  & \cross & \valid  & \warn & \cross \\
    L1 Penalty  & \valid & \valid  & \valid & \valid \\
    Lasso and elastic net  & \valid & \valid  & \warn & \cross \\
    Separating hyperplan  & \valid & \valid  & \valid & \cross \\
    MNIST classification  & \valid & \valid  & \warn & \valid \\
    Multitask Lasso  & \valid & \valid  & \warn & \valid \\
    Orthogonal matching & \valid & \valid  & \valid & \valid \\
    PCA decomposition  & \valid & \valid  & \valid & \valid \\
    Robust Linear Regression  & \valid & \valid  & \warn & \cross \\
    Segmentation toy  & \valid & \valid  & \warn & \cross \\
    SVM  & \valid & \valid  & \valid & \cross \\
    Tomography  & \valid & \cross  & \warn & \cross \\
    Weighted samples  & \valid & \valid  & \valid & \cross \\
    \hline
    \end{tabular}
    \caption{\pytracer execution summary on Scikit-learn tests. \\RR: Random Rounding. trace: outcome of \pytracer tracing, parse: outcome of \pytracer trace parsing and aggregation. \valid: successful run, \warn: error in fuzzy noise injection, \cross: numerical error or divergent traces.}
    \label{tab:pytracer_sklearn_results_summary}
\end{table}

@@

\pytracer can only aggregate traces obtained from the same control flow path. In particular, program branches triggered by an unstable floating-point comparison lead to different control flows and result in parsing errors. 
For example, comparing a residual error or the distance between two points to a strict threshold in an iterative scheme generally results in unstable branches. 
Full MCA even perturbs exact values such as integers represented with floating-point values, leading to many execution failures. For instance, the \texttt{fft1} test from SciPy raised an \texttt{ValueError: operands could not be broadcast together with shapes (601,) (600,)} error due to an array shape mismatch. A closer inspection showed that one of the arrays involved in the error was created with the NumPy function \texttt{linspace} that itself calls function \texttt{arange} (Listing~\ref{fig:pyarray_range_code}). 
Lines 3174-3178 show that the function stores the array size in a floating-point value that will be perturbed by Full MCA, as shown in table~\ref{tab:mca_result_linspace}, resulting in array sizes that fluctuate between N and N+1.
It is worth noting that 78\% of execution issues encountered in Full MCA mode without \texttt{ufunc} instrumentation
are due to this type of ``typing" errors. While they do not lead to any numerical issue, typing errors should still be fixed to preserve code quality \tristan{do we really mean that given your answer to my comment below?}. We provide solutions to help MCA better work with those cases in the discussion section.

@@

\subsubsection{Overall numerical quality}

Table~\ref{tab:pytracer_test_precision_summary} summarizes the precision for main outputs of SciPy and scikit-learn tests. 
We measured the precision in significant bits on the main test output by taking the element-wise mean for non-scalar values and, if the test produces many outputs, the average value among all of them. The numerical quality across all tests varies from 0 to 52, with an average of 44 significant bits. For both modes, we note that only one test failed due to a numerical error \tristan{can we see that in the table? if not, don't comment on it, it is puzzling. Could you show that in the table by using the same error symbols as in Table 2?} (\texttt{face\_recognition} for RR and \textit{Bayesian Ridge Regression} for Full MCA) even though we underestimate this number since Full MCA mode is more intrusive than RR mode and most of the tests produce error caused by the fuzzy noise injection.

>>>>

\documentclass[11pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{todonotes}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage{minted} % Provides listing environment
\usepackage[htt]{hyphenat}
\usepackage{subcaption}
% \usepackage{subfigure}

@@

\newcommand{\Yohan}[1]{\color{green!75!black}\textbf{Yohan:} #1\color{black}\xspace}
\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\scross}[0]{\ding{53}~}
\newcommand{\svalid}[0]{\ding{51}~}
\newcommand{\swarn}[0]{?~}
\newcommand{\pytracer}[0]{PyTracer\xspace}

\include{href}

@@

\begin{abstract}
Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. We developed \pytracer, a profiler to quantify numerical instability in Python applications to address this issue. \pytracer automatically instruments Python code at a low cost ($\simeq3.5\times$) to produce numerical traces and visualize them interactively in a Plotly dashboard. We designed \pytracer to work with any numerical noise model, including Monte-Carlo Arithmetic, random rounding or, random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. \pytracer permitted pinpointing numerical instabilities caused by a LAPACK library function and fixing them.
% Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. To address this issue, we developed \pytracer, a profiler to quantify numerical instability in Python applications. \pytracer automatically instruments Python code to produce numerical traces and visualize them interactively in a Plotly dashboard.
% \pytracer is designed to work with any type of numerical noise model, including Monte-Carlo Arithmetic, random rounding or random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. Finally, \pytracer slows down executions by only $3.5 \times$ on average.
\tristan{Add one sentence to summarize the findings} 
\end{abstract}

@@

Dynamic Binary Instrumentation (DBI) operates directly on executables, without the need for recompilation or manual changes. Therefore, it is applicable to any programming language.
CraftHPC~\cite{lam2013dynamic} uses DBI to detect catastrophic cancellations at runtime.
Verrou~\cite{fevotte2016verrou} is a Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts. FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by using shadow computations with higher precision.
Herbgring~\cite{sanchez2017finding} is a Valgrind-based tool to detect
numerical instabilities. It uses a shadow memory to detect precision losses by comparison with results obtained at higher precision. It is combined with symbolic computation to backtrack the root of the error.
We can note that all methods based on a high-precision oracle suppose that computing with a large number of bits is sufficient to obtain an accurate result, which is not always true (see for example, the famous Muller's sequence~\cite{bajard1996introduction}. %\tristan{it looks like your veritracer paper is the first time where the ``famous Muller's sequence" was described, is that so?}).

@@

Three main approaches exist to detect numerical instabilities: stochastic arithmetic, uncertainty or sensitivity analysis~\cite{helton2006survey}, and random seed analysis~\cite{click2011quality}.
%\tristan{add key references for uq/sa if any}
\pytracer focuses on stochastic arithmetic since it does not require assumptions about 
the noise model contrary to the sensitivity analysis.
Moreover, stochastic arithmetic allows evaluating the numerical quality of an implementation.
%\tristan{due to...}
Stochastic arithmetic leverages randomness to estimate numerical instabilities coming from the use of floating-point representations. The main idea is to treat round-off errors 
% \tristan{only round-off errors or all numerical errors?} 
as random variables and characterize them statistically.

@@

\pytracer's \texttt{loader} class first loads the original module, then copies it as a new instance of the \texttt{ModuleType} class. It then calls the appropriate instrumentation function for each attribute depending on its type (function, class, or instances) and replaces the original module in the \texttt{sys.modules} map with the instrumented version. Finally, it updates all existing references to the original module contained in the global symbol table.
As a result, the application will transparently call the instrumented module.

@@

    Returns:
        outputs: Outputs of the wrapped function
    """
    fid, fmodule, fname = info
    function = original_function_cache.id_dict[fid]
    bind = Binding(function, *args, **kwargs)
    stack = self.backtrace()
    time = elements()
    self.inputs(time=time,
                module_name=fmodule,
                function_name=fname,
                function=function,
                args=bind.arguments,
                backtrace=stack)
    outputs = function(*bind.args, **bind.kwargs)
    self.outputs(time=time,
                 module_name=fmodule,
                 function_name=fname,
                 function=function,
                 args=outputs,
                 backtrace=stack)
    return outputs
\end{lstlisting}
    \caption{\pytracer's generic wrapper function that dumps
    the inputs and outputs of the wrapped function.}
    \label{fig:generic_wrapper}
\end{listing}

@@

We evaluated \pytracer on two widely-used Python libraries: SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference machine-learning framework that
 implements a wide range of state-of-the-art algorithms. For each library, we executed the software examples provided in their respective GitHub repositories. We traced each example five times, having activated MCA in the Python interpreter, BLAS, LAPACK, NumPy, SciPy, and scikit-learn. We repeated the experiment with two MCA modes, Random Rounding (RR) and Full MCA, resulting in 2 experiments for each example. We used the default virtual precision in Verificarlo, namely 24 bits for single-precision floating-point values and 53 bits for double-precision values, which simulates machine error. We used Python-3.8.5, NummPy-1.19.1, SciPy-1.5.4, and scikit-learn-0.23.2.
%  \tristan{Report scipy and sklearn versions used.}

@@

Table~\ref{tab:pytracer_results_summary} summarizes
\pytracer's outcome on SciPy and scikit-learn examples.
The columns \textit{trace} and \textit{parse} report on the tracing and trace aggregation steps with three possible outcomes: green for success, red for numerical error in the application, and orange for errors due to the fuzzy noise injection
that are not considered as numerical errors, like for instance index errors due to an index shift.
% \tristan{Explain what this is, in a setence here or with a subsubsection if it requires more}. 
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random Rounding mode was entirely successful for 32/40 examples. The remaining examples failed due to numerical errors in the libraries, which we discuss hereafter. The instrumentation with Full MCA is more invasive and reduced the number of successful executions to 12/40.

The column \textit{result} summarizes the precision for main outputs of SciPy and scikit-learn examples. 
We measured the precision in significant bits on the main example output by taking the element-wise mean for non-scalar values and, if the example produces many outputs, the average value among all of them. The numerical quality across all examples varies from 0 to 52, with an average of 44 significant bits. For both modes, we note that only one example failed due to a numerical error 
% \tristan{can we see that in the table? if not, don't comment on it, it is puzzling. Could you show that in the table by using the same error symbols as in Table 2?} 
(\texttt{face\_recognition} for RR and \textit{Bayesian Ridge Regression} for Full MCA) even though we underestimate this number since Full MCA mode is more intrusive than RR mode and most of the examples produce error caused by the fuzzy noise injection.

\begin{table}[]
    \centering
    \small
    \begin{subfigure}[t]{\linewidth}
    \centering
    \begin{tabular}{|ll|c|c|c|c|c|c|}
    \hline
    \multicolumn{2}{|c}{ \multirow{2}{*}{Application} } & \multicolumn{3}{|c|}{Random Rounding (RR)} & \multicolumn{3}{c|}{Full MCA} \\
    \cline{3-8}
    & & trace & parse & result & trace & parse & result \\
    \hline
    %        &               RR                 &                 MCA               \\
    %        &  numpy         &   no numpy      &      numpy      &     no numpy    \\
    %        & run   & parse  &  run   &  parse &  run   & parse  & run    & parse  \\

    \multicolumn{1}{|c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} } }
    & \discreteCosineRF &  \valid & \valid & 52 & \warn & \cross & 47 \\ 
    \multicolumn{1}{|c|}{} & \fftOneDRf & \valid & \valid & 41 & \warn  & \cross & - \\
    \multicolumn{1}{|c|}{} & \fftTwoDRf & \valid & \valid & 52 & \valid & \valid & 52 \\
    \hline 
    \multicolumn{1}{|c|}{  \multirow{5}{2em}{ \rotatebox{90}{Interpolation} } }
    & \interOneDRf & \valid & \valid & 51 & \warn & \cross &  - \\
     \multicolumn{1}{|c|}{} & \multiRf      & \valid & \valid & 51 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & \bsplineRf    & \valid & \valid &10 &  \valid & \valid & 10\\
     \multicolumn{1}{|c|}{} & \splineOneDRf & \valid & \cross & 52 & \warn  & \cross & 52 \\
     \multicolumn{1}{|c|}{} & \splineTwoDRf & \valid & \valid & 10 &  \warn  & \cross & - \\
    \hline
     \multicolumn{1}{|c|}{ \multirow{12}{2em}{ \rotatebox{90}{Optimization} } }
    & \bfgsRf & \valid & \valid & 48  & \valid & \valid & 46 \\
     \multicolumn{1}{|c|}{} & \globalRf & \valid & \cross & 17 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & | SHGO\footnote{Simplicial Homology Global Optimization} & \valid & \cross & 25 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & | Dual Annealing & \valid & \cross & 4 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & | Differential Evolution & \valid & \cross & 0 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & | Basin Hopping & \valid & \cross & 40 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & | SHGO Sobol    & \valid & \cross & 45 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & \slsqpRf  & \valid & \valid & 45 & \valid & \valid & 44 \\
     \multicolumn{1}{|c|}{} & \lsmRf    & \valid & \valid & 48 & \warn  & \cross & - \\
     \multicolumn{1}{|c|}{} & \nelderRf & \valid & \valid & 44 & \valid & \valid & 42 \\
     \multicolumn{1}{|c|}{} & \ncgRf    & \valid & \valid & 48 & \valid & \cross & 47 \\
     \multicolumn{1}{|c|}{} & \rootRf   & \valid & \cross & 52 & \warn  & \cross & 50 \\
    \multicolumn{1}{|c|}{}  & \rootLargRf      & \valid & \cross & 29 & \valid & \cross & 26 \\
     \multicolumn{1}{|c|}{} & \rootLargePredRf & \valid & \cross & 42 & \valid & \cross & 42 \\
    \multicolumn{1}{|c|}{}  & \trustExactRf & \valid & \valid & 48  & \valid & \valid & 46 \\
     \multicolumn{1}{|c|}{} & \trustTruncRf & \valid & \valid & 48  & \valid & \valid & 46 \\
    \multicolumn{1}{|c|}{}  & \trustNCGRf   & \valid & \valid & 52  & \valid & \valid & 47 \\
    \hline
%     \end{tabular}
% \end{subfigure}
%     \label{tab:pytracer_scipy_results_summary}
% % \end{table}
% % \begin{table}[]
%     \centering
% \begin{subfigure}[t]{.8\linewidth}
%     \begin{tabular}{|c|c|c|c|c|c|c|}
%     \hline
%     \multirow{2}{5em}{Application} & \multicolumn{3}{c|}{RR} & \multicolumn{3}{c|}{MCA} \\\cline{2-7}
%                                 & trace & parse & result &  trace & parse & result \\
%     \hline
    %        &               RR                 &                 MCA               \\
    %        &  numpy         &   no numpy      &      numpy      &     no numpy    \\
    %        & run   & parse  &  run   &  parse &  run   & parse  & run    & parse  \\

    \multicolumn{1}{|c|}{ \multirow{20}{2em}{ \rotatebox{90}{Scikit-learn} } }
    & \AdaboostRf   & \valid & \valid & 48 & \warn  & \cross & - \\
    \multicolumn{1}{|c|}{} & \brrRf        & \valid & \valid & 25 & \cross & \cross & - \\
    \multicolumn{1}{|c|}{} & | $1^{st}$ dataset & \valid & \valid & 25 & \cross & \cross & - \\
    \multicolumn{1}{|c|}{} & | $2^{nd}$ dataset & \valid & \valid & nan & \cross & \cross & - \\
    \multicolumn{1}{|c|}{} & \onlineClassifierComparisonRf & \valid & \valid & 20 & \warn & \cross & -  \\
    \multicolumn{1}{|c|}{} & \kmeansRf     & \valid & \valid & 52 & \valid & \cross & - \\
    \multicolumn{1}{|c|}{} & \covarianceRf & \valid & \cross & 48 & \warn  & \cross & - \\
    \multicolumn{1}{|c|}{} & \decisionRf   & \valid & \valid & 50 & \valid & \valid & 50 \\
    \multicolumn{1}{|c|}{} & \digitsRf     & \valid & \valid & 52 & \warn  & \cross & 52 \\
    \multicolumn{1}{|c|}{} & \faceRf       & \cross & \valid & -  & \warn  & \cross & - \\
    \multicolumn{1}{|c|}{} & \penaltyRf    & \valid & \valid & -  & \valid & \valid & - \\
    \multicolumn{1}{|c|}{} & \lassoRf      & \valid & \valid & 48 & \warn  & \cross & 47 \\
    \multicolumn{1}{|c|}{} & \hyperplaneRf & \valid & \valid & 3  & \valid & \cross & - \\
    \multicolumn{1}{|c|}{} & \mnistRf      & \valid & \valid & 43 & \warn  & \valid & - \\
    \multicolumn{1}{|c|}{} & \multitaskRf  & \valid & \valid & 50 & \warn  & \valid & 48 \\
    \multicolumn{1}{|c|}{} & \ompRf        & \valid & \valid & 52 & \valid & \valid & 52 \\
    \multicolumn{1}{|c|}{} & \pcaRf        & \valid & \valid & 48 & \valid & \valid & 48 \\
    \multicolumn{1}{|c|}{} & \robustRf     & \valid & \valid & 50 & \warn  & \cross & 48 \\
    \multicolumn{1}{|c|}{} & | Linear        & \valid & \valid & 48 & \warn  & \cross & 46 \\
    \multicolumn{1}{|c|}{} & | RANSAC        & \valid & \valid & 50 & \warn  & \cross & 48 \\
    \multicolumn{1}{|c|}{} & \toyRf        & \valid & \valid & 35 & \warn  & \cross & - \\
    \multicolumn{1}{|c|}{} & \svmRf        & \valid & \valid & 43.5 & \valid & \cross & - \\
    \multicolumn{1}{|c|}{} & | Non-weighted  & \valid & \valid & 43 & \valid & \cross & - \\
    \multicolumn{1}{|c|}{} & | Weighted      & \valid & \valid & 44 & \valid & \cross & - \\
    \multicolumn{1}{|c|}{} & \tomographyRf & \valid & \cross & - & \warn  & \cross & - \\
    \multicolumn{1}{|c|}{} & \weightedRf   & \valid & \valid & 44 & \valid & \cross & - \\
    \hline
    \end{tabular}
    \end{subfigure}
    \caption{\pytracer execution summary on SciPy and Scikit-learn examples. \\ \emph{trace}: outcome of \pytracer tracing, \emph{parse}: outcome of \pytracer trace parsing and aggregation, \emph{result}: precision for main outputs in significant bits. \valid: successful run, \warn: error from invalid noise injection, \cross: numerical error or divergent traces. \textit{nan} stands for (\texttt{Not-A-Number}). The names example are linked to their source code.}
    \label{tab:pytracer_results_summary}
\end{table}

    % \caption{
    % \pytracer execution summary on SciPy examples.\\RR: Random Rounding. trace: outcome of \pytracer tracing, parse: outcome of \pytracer trace parsing and aggregation.  \valid: successful run, \warn: error in fuzzy noise injection, \cross: numerical error or divergent traces.
    % }

@@

\pytracer can only aggregate traces obtained from the same control flow path. In particular, program branches triggered by an unstable floating-point comparison lead to different control flows and result in parsing errors. 
For example, comparing a residual error or the distance between two points to a strict threshold in an iterative scheme generally results in unstable branches. 
Full MCA even perturbs exact values such as integers represented with floating-point values, leading to many execution failures. For instance, the \texttt{fft1} example from SciPy raised an \texttt{ValueError: operands could not be broadcast together with shapes (601,) (600,)} error due to an array shape mismatch. A closer inspection showed that one of the arrays involved in the error was created with the NumPy function \texttt{linspace} that itself calls function \texttt{arange} (Listing~\ref{fig:pyarray_range_code}). 
Lines 3174-3178 show that the function stores the array size in a floating-point value that will be perturbed by Full MCA, as shown in Table~\ref{tab:mca_result_linspace}, resulting in array sizes that fluctuate between N and N+1.
It is worth noting that 78\% of execution issues encountered in Full MCA mode instrumentation
are due to this type of ``typing" errors.  
In conclusion, those regions of code should not be perturbed by MCA since the injected noise does not make sense.
\tristan{conclusion: ces regions du code ne devraient pas etre bruitees car le bruit n'a pas de sens ici (exemple du spin dans -1/2,1/2. Ca correspond aux ?}
% While they do not lead to any numerical issue, typing errors should still be fixed
% by improving the Full MCA mode instrumentation in order to preserve code correctness.
\tristan{do we really mean that given your answer to my comment below?}. 
\Yohan{By reading again this sentence, I agree that it is confusing. 
It suggests that we should fixed the code while my idea was to fix the MCA instrumentation.}
We provide solutions to help MCA better work with those cases in the discussion section.

@@

% \subsubsection{Overall numerical quality}
